---
title: "Reproducible Research: Peer Assessment 1"
author: Pedro A. Alonso Baigorri
date: 22/05/2017
output: 
  html_document:
    keep_md: true
---
## Introduction
This document describes the analysis done for the 1st practice of the Reproducible Research Course, part of the Data Science specialization in course.

The objectives of the exercice are described in the README.md file on this repository.


## Setting global options
```{r global_options, include=FALSE}
knitr::opts_chunk$set(fig.width=12, fig.height=8, fig.path='figure/',
                      echo=TRUE, warning=FALSE, message=FALSE)
```

## Loading and preprocessing the data
```{r}
setwd('.')
unzip("activity.zip")
steps_dataset <- read.csv("activity.csv")
head(steps_dataset)
```

## Histogram of the total number of steps taken each day
The purpose of the histogram is to know what's the frecuency (number of days) where an specific value of the total steps has occured during all the period of analysis. In summary I will represent a barplot where the y-axis is the frequency (number of days) and the x-axis the different values of total  number of steps.

To do this I will calculate first the total number of the steps by day and then plot an histogram using the hist() function.

```{r}
steps_by_day <- aggregate(steps ~ date, data = steps_dataset, FUN = sum )
head(steps_by_day)
hist(steps_by_day$steps, breaks=30, col=c("red"), xlab = "Number of steps", main =" Histogram of the total number of steps taken each day")

```

## What is the mean and median of  total number of steps taken per day?

Using the agreggated data calculated before, the mean and the median can be calculated as follows:

```{r results="hide"}
mean_steps <- mean(steps_by_day$steps)
median_steps <- median(steps_by_day$steps)
```

This provides the following results:  
  
- Mean = `r mean_steps`  
- Median = `r median_steps`


## What is the average daily activity pattern?
To analyze the average daily activity pattern first of all I will make a time series plot of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis)

```{r}
steps_by_interval <- aggregate(steps ~ interval, data = steps_dataset, FUN = mean )
head(steps_by_interval)
plot(steps_by_interval$interval, steps_by_interval$steps,  ylab = "Average of steps", xlab = "interval", type = "l", main = "Daily activity pattern")
```

Now I'll calculate the maximum value and  interval when this maximum number happens
```{r results = "hide"}
max_steps <- max(steps_by_interval$steps)

id_max_interval <- which(steps_by_interval$steps == max_steps)
max_interval <- steps_by_interval[max(id_max_interval), "interval"]

#another way to calculate the max_interval
max_interval <- steps_by_interval[steps_by_interval$steps==max_steps,"interval"]
```

The max avg steps are: `r max_steps`.  
The interval when the max avg step happens is: `r max_interval`





## Imputing missing values



## Are there differences in activity patterns between weekdays and weekends?
